RTS AI: Problems and Techniques

### 2015-12-14 ECGG_RTSProblems/ECGG_RTSProblems.tex
Fix some typos and AIIDE references

\section*{Introduction}\label{sec:intro}
The  field of  Real-Time Strategy  (RTS) game  Artificial Intelligence
(AI) has advanced significantly in the past few years, partially thanks to competitions as the ``ORTS
RTS  Game AI  Competition''  (held  from 2006  to  2009), the  ``AIIDE
StarCraft AI Competition'' (held since  2010), and the ``CIG StarCraft
RTS AI Competition'' (held since  2011). Based on the work work presented in \cite{ontanon2013survey}, here we first define RTS games, then list the open problems in creating AI for RTS games, and finally point to the approaches that have been proposed to address these problems.

@@

\subsubsection*{Uncertainty}
Adversarial planning under  uncertainty in domains of the  size of RTS
games is  still an unsolved  challenge.  In  RTS games, there  are two
main kinds  of uncertainty. First,  the game is  partially observable,
and players  cannot observe the  whole game  map (like in  Chess), but
need to scout in order to see what the opponent is doing. This type of
uncertainty   can  be   lowered  by   good  scouting,   and  knowledge
representation  (to  infer  what  is  possible  given  what  has  been
seen). However, scouting also means deliberately reducing economic progress in order to obtain information.
Second,  there is also  uncertainty arising from the  fact that
the games  are adversarial,  and a player  cannot predict  the actions
that the opponent(s)  will execute. For this type  of uncertainty, the
AI, as the human  player, can only build a sensible  model of what the
opponent is likely to do.
% [Santi:  I  commented  out  this  part,  since  I'm  not  sure  it's
% correct.  The second  type  of  uncertainty can  also  be linked  to
% data.  For example,  by analyzing  player replays,  one could  learn
% "plan recognition"  probability tables, which  can be used  to infer
% the strategy  of the opponent,  and thus, reduce the  uncertainty of
% the opponent's future  actions] Both these kinds  of uncertainty can
% be studied through probabilities, but only the first type, $P(Hidden
% | Seen)$, can be {\em easily} linked to
% data. %now for the clumsy part:
% One could  argue that the  extensional data  is the only  thing that
% matters because  it is the only  thing that has an  influence on the
% game, but in  reality, for very highly  skilled players, discovering
% the intentions of the  enemy is the key to winning.  It allows for a
% compact understanding of the game  state and development, while also
% giving additional information about the  future. Whereas a player is
% doing action  A to  follow it  up by action  B or  C depends  on its
% intention,  while the  information  about action  A  is visible  and
% information about  B or  C is  hidden, only  intention allows  for a
% choice.

@@

% INTRO:
%Tactical reasoning involves reasoning about the different abilities of
%the units in a group and about the environment (terrain) and positions
%of the different  groups of units in order to  gain military advantage
%in battles. For  example, it would be a very  bad tactical decision to
%send  fast, invisible  or flying  units (typically  expensive) in  the
%first line  of fire against slower  heavier units, since they  will be
%wiped out fast.  
We will divide the work on mid-range tactical reasoning in RTS games in two
large groups: spatial reasoning and decision making (that has been addressed both using machine learning and game tree search).

>>>>

\section*{Introduction}\label{sec:intro}
The  field of  Real-Time Strategy  (RTS) game  Artificial Intelligence
(AI) has advanced significantly in the past few years, partially thanks to competitions as the ``ORTS
RTS  Game AI  Competition''  (held  from 2006  to  2009), the  ``AIIDE
StarCraft AI Competition'' (held since  2010), and the ``CIG StarCraft
RTS AI Competition'' (held since  2011). Based on the work presented in \cite{ontanon2013survey}, here we first define RTS games, then list the open problems in creating AI for RTS games, and finally point to the approaches that have been proposed to address these problems.

@@

\subsubsection*{Uncertainty}
Adversarial planning under  uncertainty in domains of the  size of RTS
games is  still an unsolved  challenge.  In  RTS games, there  are two
main kinds  of uncertainty. First,  the game is  partially observable,
and players  cannot observe the  whole game  map like in  Chess, but
need to scout in order to see what the opponent is doing. This type of
uncertainty   can  be   lowered  by   good  scouting,   and  knowledge
representation  (to  infer  what  is  possible  given  what  has  been
seen). However, scouting also means deliberately reducing economic progress in order to obtain information.
Second,  there is also  uncertainty arising from the  fact that
the games  are adversarial,  and a player  cannot predict  the actions
that the opponent(s)  will execute. For this type  of uncertainty, the
AI, as the human  player, can only build a sensible  model of what the
opponent is likely to do.
% [Santi:  I  commented  out  this  part,  since  I'm  not  sure  it's
% correct.  The second  type  of  uncertainty can  also  be linked  to
% data.  For example,  by analyzing  player replays,  one could  learn
% "plan recognition"  probability tables, which  can be used  to infer
% the strategy  of the opponent,  and thus, reduce the  uncertainty of
% the opponent's future  actions] Both these kinds  of uncertainty can
% be studied through probabilities, but only the first type, $P(Hidden
% | Seen)$, can be {\em easily} linked to
% data. %now for the clumsy part:
% One could  argue that the  extensional data  is the only  thing that
% matters because  it is the only  thing that has an  influence on the
% game, but in  reality, for very highly  skilled players, discovering
% the intentions of the  enemy is the key to winning.  It allows for a
% compact understanding of the game  state and development, while also
% giving additional information about the  future. Whereas a player is
% doing action  A to  follow it  up by action  B or  C depends  on its
% intention,  while the  information  about action  A  is visible  and
% information about  B or  C is  hidden, only  intention allows  for a
% choice.

@@

% INTRO:
%Tactical reasoning involves reasoning about the different abilities of
%the units in a group and about the environment (terrain) and positions
%of the different  groups of units in order to  gain military advantage
%in battles. For  example, it would be a very  bad tactical decision to
%send  fast, invisible  or flying  units (typically  expensive) in  the
%first line  of fire against slower  heavier units, since they  will be
%wiped out fast.  
We will divide the work on mid-range tactical reasoning in RTS games in two
large groups: spatial reasoning and decision making. That has been addressed both using machine learning and game tree search.
